---
title: "strings_and_factors"

output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(rvest)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%")
theme_set(theme_minimal()+theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Strings and regex

```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")
```

```{r}
str_replace(string_vec, "jeff", "Jeff")
```

For exact matches, you can designate matches at the beginning or end of a line.

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")
```

```{r}
str_detect(string_vec, "i think$")
```

You can designate a list of characters that will count as a match.

```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Pp]umpkin")
```

You don’t have to list these; instead, you can provide a range of letters or numbers that count as a match.

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]")
```

The character `.` matches anything.

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

Some characters are “special”. These include ` [ `and` ]`, `( `and `)`, and `..` If you want to search for these, you have to indicate they’re special using` \`. Unfortunately, `\ `is also special, so things get weird.

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```

## Thoughts on factors

```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex
```

```{r}
as.numeric(vec_sex)
```

```{r}
vec_sex = fct_relevel(vec_sex, "male")
vec_sex
```

```{r}
as.numeric(vec_sex)
```

## NSDUH

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) |> 
  html_table() |> 
  first() |>
  slice(-1)
```

```{r}
data_marj = 
  table_marj |>
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))

```

## NSDUH -- factors

```{r}
data_marj |>
  filter(age == "12-17") |> 
  mutate(State = fct_reorder(State, percent)) |> 
  ggplot(aes(x = State, y = percent, color = year)) + 
    geom_point() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
## Restaurant inspections

```{r}
data("rest_inspec")

rest_inspec |> 
  group_by(boro, grade) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = grade, values_from = n)
```
To simplify things, I’ll remove inspections with scores other than A, B, or C, and also remove the restaurants with missing boro information. I’ll also clean up boro names a bit.

```{r}
rest_inspec =
  rest_inspec |>
  filter(grade %in% c("A", "B", "C"), boro != "Missing") |> 
  mutate(boro = str_to_title(boro))
```

Let’s focus only on pizza places for now, and re-examine grades by borough.

```{r}
rest_inspec |> 
  filter(str_detect(dba, "Pizza")) |> 
  group_by(boro, grade) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = grade, values_from = n)
```

That doesn’t look right – for sure there are more pizza place ratings than that! The problem is that the match in str_detect is case-sensitive until we tell it not to be:

```{r}
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |> 
  group_by(boro, grade) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = grade, values_from = n)
```

The table is okay but I’d like to visualize this.

```{r}
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |>
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 
```
Might help to have things in a different order – maybe number of pizza places? Fortunately this can be done using `fct_infreq`.

```{r}
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |>
  mutate(boro = fct_infreq(boro)) |>
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 
```

Suppose I want to rename a borough. I could try using str_replace.

```{r}
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |>
  mutate(
    boro = fct_infreq(boro),
    boro = str_replace(boro, "Manhattan", "The City")) |>
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 
```

If I tried base R (maybe because I found some code online), I might try to do this using `replace`.

```{r}
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |>
  mutate(
    boro = fct_infreq(boro),
    boro = replace(boro, which(boro == "Manhattan"), "The City")) |>
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```
Fortunately there is a dedicated function for renaming factor levels:

```{r}
rest_inspec |> 
  filter(str_detect(dba, regex("pizza", ignore_case = TRUE))) |>
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro, "The City" = "Manhattan")) |>
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()
```



